A Hierarchical Bayesian Framework for Inferring the Stellar Obliquity Distribution

### 2023-03-23 src/tex/ms.tex
[showyourwork] overleaf sync

Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system. In transiting exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured. To determine the true stellar obliquity $\psi$, information about the stellar inclination $i_\star$ is also needed.
In this paper, we demonstrate that while a constraint on the stellar inclination is crucial for measuring the obliquity of an individual system, it is not required for robust determination of the population-level stellar obliquity distribution. 
In practice, the constraints on the stellar obliquity distribution are mainly driven by the sky-projected stellar obliquities.
We introduce a flexible, hierarchical Bayesian framework that can be used to infer the stellar obliquity distribution solely from sky-projected stellar obliquities, including stellar inclination measurements when available.
When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees. Misaligned systems have nearly isotropic stellar obliquities, which could have important implications for the formation and evolution of close-in planets.

@@

In this section, we find the expression of $\lambda$ and $i_\star$ in terms of $\psi$, $\theta$, and $i_{\rm orb}$. To do so, we introduce two coordinates that describe the stellar spin axis and its planet's orbital axis.

>>>>

Stellar obliquity $\psi$, the angle between a planet's orbital axis and its host star's spin axis, traces the formation and evolution of a planetary system. In transiting exoplanet observations, only the sky-projected stellar obliquity $\lambda$ can be measured. To determine the true stellar obliquity $\psi$, information about the stellar inclination $i_\star$ is also needed.
In this paper, we demonstrate that while a constraint on the stellar inclination is crucial for measuring the obliquity of an individual system, it is not required for robust determination of the population-level stellar obliquity distribution. 
In practice, the constraints on the stellar obliquity distribution are mainly driven by the sky-projected stellar obliquities.
We introduce a flexible, hierarchical Bayesian framework that can be used to infer the stellar obliquity distribution solely from sky-projected stellar obliquities, including stellar inclination measurements when available.
When applying the framework to all exoplanetary systems with measured sky-projected stellar obliquity, we find that the inferred population-level obliquity distribution is unimodal and peaked at zero degrees. Misaligned systems have nearly isotropic stellar obliquities \edits{with no strong clustering near 90$\degr$. Particularly, the misaligned Hot Jupiter population presents a broad stellar obliquity distribution, which may suggest planet-planet scattering after a convergent disk migration or planetary Kozai might be the dominant mechanisms sculpting the population.}

@@

In this section, we find the expression of $\lambda$ or $i_\star$ in terms of $\psi$, $\theta$, and $i_{\rm orb}$. To do so, we introduce two coordinate systems that describe the stellar spin axis $\vb{n_\star}$ and its planet's orbital axis $\vb{n_{\rm orb}}$, shown in Figure~\ref{fig:coord}.
